---
title: "Data Science for Public Policy"
author: "Nathalie Kirsch and Vanaaisha Pamnani"
output: html_document
urlcolor: blue
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r include = FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

### PPOL 670 | Assignment 4
### Data Visualization, Markdown, and Git]

[72 Small and Medium-Size Cities Built Environment and Health Equity Dataset](https://datacatalog.urban.org/dataset/72-small-and-medium-size-cities-built-environment-and-health-equity-dataset)

This Urban Institute data set includes summary demographics, economic and 
housing demographics, andhealth indicators for 72 small and medium-size cities
across the country. The research team interviewedstakeholders from 19 of the 
cities included in the data set. The remaining 53 cities were selected as “peer”
cities that shared one for more of the following characteristics with many of 
the 19 cities: similar city type (suburban, central city, etc.), range of 
population (small or medium-sized), and location within the samemetropolitan
region or the same state/part of the country

```{r}
library(tidyverse)
library(ggplot2)
library(readxl)
getwd()
```


```{r}
data <- read_xlsx(path = "data/Built Environment Health Equity Data Table.xlsx", sheet = "City data", range = "A2:AU72") %>%
  janitor::clean_names()

glimpse(data)
```

```{r}




```



